---
title: "BIOS226 Topic 5 — Parameter Exploration Worksheet"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

# Purpose

Use `run_tumor_model_quick.R` to explore how **data generation** and **model parameters** change classification performance.

This worksheet is designed for hands-on experimentation and short written interpretation.

# Before You Start

1. Open `run_tumor_model_quick.R` in your **Topic5** project folder.
2. Set both seeds to your own student ID:
   - `data_seed <- <your_student_id>`
   - `model_seed <- <your_student_id>`
3. Keep your student ID seeds for all runs unless a question explicitly asks you to change them.

Using your own student ID as seed makes your dataset and split reproducible and unique to you.

# Quick Run Pattern

For each question:

1. Edit parameters in `run_tumor_model_quick.R`.
2. Re-run:

```{r eval=FALSE}
source("run_tumor_model_quick.R")
```

3. Read results from:
   - ROC plot (test curve, random line, CV envelope)
   - truth table plot
   - `results` object in the Environment

Optional helper chunk to extract metrics:

```{r eval=FALSE}
safe_ratio <- function(num, den) if (den == 0) NA_real_ else num / den
tp <- results$test$tp
tn <- results$test$tn
fp <- results$test$fp
fn <- results$test$fn

sensitivity <- safe_ratio(tp, tp + fn)
precision <- safe_ratio(tp, tp + fp)
accuracy <- safe_ratio(tp + tn, tp + tn + fp + fn)

data.frame(
  test_auc = results$test$auc,
  tp = tp, tn = tn, fp = fp, fn = fn,
  sensitivity = sensitivity,
  precision = precision,
  accuracy = accuracy
)
```

# Questions (Answer All 10)

## Q1. Baseline with your student ID seeds

Run with default parameters (except replacing seeds with your student ID).  
Report:

- Test AUC
- TP, TN, FP, FN
- Sensitivity, Precision, Accuracy

Briefly describe the ROC shape and where the test curve sits relative to the random line.

## Q2. Threshold trade-off

Keep all settings fixed. Run three thresholds: `0.30`, `0.50`, `0.80`.

For each threshold, record TP/TN/FP/FN, Sensitivity, and Precision.  
Which threshold gives the best balance for this dataset, and why?

## Q3. Class imbalance

Keep seeds fixed. Compare:

- `class_proportion <- 0.5`
- `class_proportion <- 0.8`

How do confusion matrix counts and Precision/Sensitivity change?  
Is AUC affected as strongly as threshold-based metrics?

## Q4. Noise level

Keep everything else fixed and compare:

- `noise_sd <- 0.8`
- `noise_sd <- 2.0`

What happens to ROC AUC and error types (FP/FN) as noise increases?

## Q5. Number of informative genes

Keep `n_genes` fixed and compare:

- `n_informative <- 10`
- `n_informative <- 100`

How does increasing true signal features affect model discrimination?

## Q6. Sample size

Keep class proportion and noise fixed. Compare:

- `n_samples <- 80`
- `n_samples <- 240`

How do metric stability and ROC smoothness change with sample size?

## Q7. Feature dimensionality (p >> n stress)

Keep `n_informative` fixed and compare:

- `n_genes <- 2000`
- `n_genes <- 20000`

Does high dimensionality hurt test performance under this workflow, or is performance mostly preserved?

## Q8. Top-k selected genes

Keep data settings fixed. Compare:

- `top_k_genes <- 10`
- `top_k_genes <- 25`
- `top_k_genes <- 100`

Which value performs best for your student-ID dataset?  
Give one reason based on bias/variance or signal-to-noise thinking.

## Q9. Number of CV folds

Keep everything else fixed. Compare:

- `k_folds <- 3`
- `k_folds <- 5`
- `k_folds <- 10`

Inspect the CV envelope on the ROC plot.  
How does fold choice affect the apparent stability of CV performance?

## Q10. Reproducibility and personalization check

1. Re-run once with exactly the same settings and your student ID seeds.  
2. Then change both seeds to `student_id + 1` and re-run.

What stayed identical in run 1 vs run 2, and what changed after updating seeds?  
Explain in 2-4 sentences why seed control is important in model comparison.

# Submission Prompt

In your submission, include:

- A short table of your key runs (parameter changes + metrics).
- Direct answers to Q1-Q10.
- One final paragraph: which parameter had the largest practical impact on your model quality, and why?
